---
title: "Hierarchical Clustering of Santa Barbara Stream Chemistry"
description: |
  Hierarchical clustering by complete linkage to create a dendrogram showing multivariate clustering for water chemistry by site in Santa Barbara, CA.
author:
  - name: Catherine Takata
    url: {}
date: 03-14-2021
output:
  distill::distill_article:
    self_contained: false
---

### Data Source: 
SBC LTER: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000
Creators: Santa Barbara Coastal LTER, & Melack, John M
Citation: Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. Environmental Data Initiative. https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174. 

### Load in Packages

```{r setup}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)

# Packages for cluster analysis:
library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)
```

Create a dataset of the specified variables of the Santa Barbara Stream Chemistry data. Use tidy data techniques and reassign -999 values as "NA". 

```{r}
# Read in data and convert -999 values to NA.
water_chem <- read.csv(here("data", "sbc_lter_registered_stream_chemistry.csv")) %>% 
  na_if(-999.0) %>% 
  clean_names()
```

Average the observations across all columns in order to obtain one value per site. 

```{r}
# Take means from all water chem variables (observations) per site
water_chem$summary <- rowMeans(water_chem[3:12], na.rm = TRUE)

# Create a data set that has a single summary row per site 
water_sum <- water_chem %>% 
  dplyr::select(site_code, summary) %>% 
  group_by(site_code) %>% 
  summarise(mean = mean(summary))
```

```{r}
# Scale the numeric variables 
water_scaled <- water_sum %>% 
  dplyr::select(2) %>% 
  scale()
  
# Update to add rownames (site_code) from water_chem
rownames(water_scaled) <- water_sum$site_code

```

### Euclidean distances
Use the `stats::dist()` function to find the Euclidean distance in multivariate space between the different water quality indexes (by site):

```{r}
# Compute dissimilarity values (Euclidean distances):
euc_distance <- dist(water_scaled, method = "euclidean")
```

### Complete Linkage Dendrogram
Use stats::hclust() function to calcualte hierarchical clustering based on a matrix of euclidean distances. 
Use complete linkages - merge clusters by the smallest maximum distance between two observations in distinct clusters. 

```{r, fig.show = 'hide'}
# Hierarchical clustering (complete linkage)
cluster_complete <- hclust(euc_distance, method = "complete" )

# Plot it (base plot):
plot(cluster_complete, cex = 0.6, hang = -1)
```

```{r}
ggdendrogram(cluster_complete, 
             rotate = TRUE) +
  theme_minimal() +
  labs(title = "Hierarchical Clustering of Santa Barbara Stream Chemistry by Site",
       x = "Sites",
       y = "")
```

