---
title: "California Cetacean Species Spatial Extent"
description: |
  Create a raster of species richness of 35 cetacean species off the coast of California displayed in a static map.
author:
  - name: Catherine Takata
    url: {}
date: 03-14-2021
output:
  distill::distill_article:
    self_contained: false
---

### Download data from Aquamaps
- Download the file of rasters from AquaMaps (Kaschner, K., Rius-Barile, J., Kesner-Reyes, K., Garilao, C., Kullander, S., Rees, T., & Froese, R. (2016). AquaMaps: Predicted range maps for aquatic species. www.aquamaps.org),


### Load In Packages

```{r setup, include = T}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(raster) 
library(tidyverse)
library(here)
library(sf)
library(fasterize)
library(rnaturalearth)
```


```{r, include=T}
# Load in subsetted raster data 
# Use rasterstack() to read in multiple layers at once as a multi-layer 
ca_species <- list.files(path = here("data","ca_cetaceans"), pattern = "*.tif", full.names = TRUE)
species_stack <- raster::stack(ca_species)
```

With probability of occurrence data such as AquaMaps, it is common to decide on a threshold value above which a species is considered “present.” Choose a probability threshold and use your threshold to reclassify each probability layer to a “presence” layer of either 0 (not present) or 1 (present). Here, we will set our threshold = 0.60.


```{r, include=T}
# Set a probability threshold using a if else statement to reclassify the presence of species 
# If the species is greater than or equal to a 0.6 threshold, return 1 ~ "present" else return 0 ~ "absent" 
cetaceans_thresh <- function(x, thresh = .6) {
  y <- ifelse(x >= thresh, 1, NA)
  return(y)
}

# Use the calc() function to determine the species richness of the compounded raster stack 
# Determines how many species are "present" in each cell 
species_richness <- calc(species_stack, fun = cetaceans_thresh)
species_richness_sum <- calc(species_richness, fun = sum, na.rm = TRUE)
```


Convert the raster data to a dataframe to plot with ggplot().  The `raster::rasterToPoints()` function converts a raster to a matrix of points - `x`, `y`, and `layer` (or the layer name if it's a named layer).  Then `as.data.frame()` can convert that matrix to a `data.frame` object suitable for `ggplot.`


```{r, include=T}
species_richness_df <- raster::rasterToPoints(species_richness) %>%
  as.data.frame()
species_sum_df <- raster::rasterToPoints(species_richness_sum) %>%
  as.data.frame()
```


```{r, message=F}
# Using rnaturalearth, obtain a pre-downloaded vector map for coastline data 
# Coastline data will add visual context to our raster data 
coastline <- ne_download(scale = 50, type = 'coastline', category = 'physical', returnclass = "sf")

# Use ggplot() to plot and populate with raster data + spatial data to produce a "map" 
ggplot() + 
  geom_raster(data = species_sum_df, aes(x = x, y = y, fill = layer)) +
  geom_sf(data = coastline) +
  coord_sf(xlim = c(-125, -115), ylim = c(32, 38)) + 
  scale_fill_gradient(low = 'white', high = 'blue') +
  theme_void() +
  labs(title = "Species richness of cetacean species of CA",
       caption = "Figure 1")
```

**Figure 1**: Red shading indicates species richness of 35 cetacean species off the California coast. Probability of occurence was determined by choosing a probability threshold of 0.6 to indicate "presence" of each species. Increased saturation indicates higher density of species and more occurences. 


Building raster data can be tricky. Since each raster contains data, often cells overlap and merge. Obtaining higher resolutuon would demand more data. For now, thsi is a good start at visualizing species richness of CA cetacean species. 
