---
title: "Visualizing California Oil Spills"
description: |
  Using various spatial data wrangling and visualization methods, we will explore the California oil spill incidents gathered by the The Office of Spill Prevention and Response (OSPR) Incident Tracking Database." 
  
author:
  - name: Catherine Takata
date: 03-14-2021
output:
  distill::distill_article:
    self_contained: false
---

### Download data from CA DFW Oil Spill Incident Tracking: 
- https://map.dfg.ca.gov/metadata/ds0394.html
- https://gis.data.ca.gov/datasets/7464e3d6f4924b50ad06e5a553d71086_0/data

Source: Oil Spill Incident Tracking [ds394], 2009-07-23"

### Load In Packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(janitor)
library(here)
library(broom)

# Spatial data packages
library(sf)
library(tmap)
```

Establish a new dataset based on specific parameters of the counties of California. Here, we will "read in" the data to isolate specific areas and their regional extent. This will provide a spatial basis that will be populated with designated points. 


```{r, warning = F}
# Read in California county shapefile 
# Shapefile will act as the polygon needed for visualization/population of data 
# Data wrangle to manipulate data easily, isolate two attributes using select() and rename() for better recall 

ca_counties <- read_sf(here("data", "ca_counties", "CA_Counties_TIGER2016.shp")) %>% 
  select(NAME, ALAND) %>% 
  rename(county_name = NAME, land_area = ALAND)
```

Read in the "oil spill" data. This dataset will overlay the map parameters we established earlier. 

```{r, warning = F}
# Read in CA DFW Oil Spill dataset ds394
oil_spill <- read_sf(here("data", "Oil_Spill_Incident_Tracking_%5Bds394%5D-shp")) %>% 
  filter(INLANDMARI == "Inland")
```


### Exploratory interactive map 
Use {tmap} to make an interactive map. We can explore the extent of the data, zoom in, and zoom out. 


```{r, warning = F, echo = T}
tmap_mode(mode = "view")

tm_shape(ca_counties) +
  tm_fill("land_area")
```

Where do we see the highest number of spills? To gain some more insight, let's create a chloropleth map! 

We can identify the density of oil spills per area. With similar data wrangling techniques, we can specify new paramters. 

```{r, warning = F}
# Use st_join() to join the datasets to obtain counts by county 
ca_oil_spill <- ca_counties %>% 
  st_join(oil_spill)

# Use counts() to find counts of incidents by county 
ca_oil_spill_counts <- ca_oil_spill %>% 
  count(county_name)
```

### Finalized chloropleth map 

```{r, echo = T}
# Plot a chloropleth using the number of oil spill incidents as the fill color 

ggplot(data = ca_oil_spill_counts) + 
  geom_sf(aes(fill = n), color = "white", size = 0.1) + 
  scale_fill_gradientn(colors = c("lightgray", "orange", "red")) + 
  theme_bw() + 
  labs(title = "CA inland oil spill incidents by county (2008)",
       fill = "Number of oil spills")

```







